### 2021-01-22 paper/arm.tex
Initial definitions.
More or less feature complete in the paper.

@@

Internal visibility:
If $\bEv\xpoloc\aEv$ then either
\begin{itemize}
\item %$\bEv\xca\aEv$:
  $\bEv:\DWP{x}{1}\xco\aEv:\DWP{x}{2}$ 
\item %$\bEv\xrfx\aEv$:
  $\bEv:\DWP{x}{1}\xrfx\aEv:\DRP{x}{1}$
\item %$\bEv\mathrel{\xco\xrfx}\aEv$:
  $\bEv:\DWP{x}{1}\xco\DWP{x}{2}\xrfx\aEv:\DRP{x}{2}$
\item
  $\bEv:\DRP{x}{0}\xfr\aEv:\DWP{x}{1}$
\item %$\bEv\mathrel{\xfr\xrfx}\aEv$:
  $\bEv:\DRP{x}{0}\xfr\DWP{x}{1}\xrfx\aEv:\DRP{x}{1}$
\item %$\bEv\mathrel{\xrfxinv\xrfx}\aEv$:
  $\bEv:\DRP{x}{1}\xrfxinv\DWP{x}{1}\xrfx\aEv:\DRP{x}{1}$
  % \begin{tikzinline}
  %   \node(a){$\DWP{x}{1}$};
  %   \node(b)[right=1.5em of a]{$\DRP{x}{1}$};
  %   \node(c)[right=1.5em of b]{$\DRP{x}{1}$};
  %   \rfx{a}{b}
  %   \rfx[out=15,in=165]{a}{c}
  % \end{tikzinline}
\end{itemize}
If $\bEv\xpoloc\aEv$ then the following are impossible:
\begin{itemize}
\item $\bEv:\DWP{x}{2}\xcoinv\aEv:\DWP{x}{1}$ 
\item $\bEv:\DWP{x}{1}\xfrinv\aEv:\DRP{x}{0}$ 
\item $\bEv:\DRP{x}{1}\xrfxinv\aEv:\DWP{x}{1}$ 
\item $\bEv:\DRP{x}{2}\xrfxinv\DWP{x}{2}\xcoinv\aEv:\DWP{x}{1}$
\item $\bEv:\DRP{x}{1}\xrfxinv\DWP{x}{1}\xfrinv\aEv:\DRP{x}{0}$
\end{itemize}

@@

An ARM execution graph $G$ defines many relations, including program order
($\rpox$), reads-from ($\rrfx$), coherence ($\rco$) and several dependency
orders.  From these are derived:
\begin{itemize}
\item ${\rpoloc}$, which is the subrelation of $\rpox$ that only relates
  actions on the same location,
\item ${\rob}$, which is required to be acyclic (\ref{external}), and
\item $\reco$, with the requirement that ${\rpoloc}\cup{\reco}$ be acyclic (\ref{sc-per-loc}).
\end{itemize}
% Let $G$ be an execution graph satisfying the ARM consistency 
%requirements.
Given an execution graph $G$, we say that $\aEv$ is an \emph{internal read} if $\aEv\in\fcodom({\rpox}\cap {\rrfx})$.

@@

% Use these to refer to the rules in text:
%\ref{rf-comp} 
%\ref{co-tot}
%\ref{sc-per-loc}
%\ref{external}

@@

From $G$ we construct a candidate pomset $\aPS$ as follows:
\begin{itemize}
\item $\Event= \textsf{E}$,
\item $\labelingAct(\aEv)=\tau \mathsf{lab}(e)$, if $\aEv$ is a relaxed
  internal read, 
\item $\labelingAct(\aEv)=\mathsf{lab}(e)$, if $\aEv$ is not a relaxed
  internal read,
\item $\labelingForm(\aEv)=\TRUE$,
%\item ${\le} = {\rob}$, and
\item ${\gtN} = ({\rob} \cup {\reco})^*$
\end{itemize}
To reempphasize, in this candidate pomset, $\rob$ is calculated by considering the definition of $\rob$ without $\rrfi$, ie.:
\begin{align*}
  \tag{Dependency order}
  {\rdob} &\eqdef\smash{
    ({\raddr}\cup{\rdata});
    \cup ({\rctrl}\cup{\rdata}); {\IDW}; {\rcoi}^?
    \cup {\raddr}; {\rpox}; {\IDW}
  }
  \\
  \tag{Barrier order}
  {\rbob} &\eqdef\smash{
    {\IDAcq}; {\rpox}
    \cup {\rpox};{\IDRel}; {\rcoi}^?
  }
  \\
  \tag{Acyclic order}
  {\rob} &\eqdef\smash{
    ({\robs} \cup {\rdob} \cup {\rbob})^+
  }
\end{align*}

>>>>

\subsection{Down-Grading Acquires ($\xDGA$)}
\label{sec:internal}

We use the uninterpreted symbols $\Dx{\aLoc}$.

\begin{definition}
  \label{def:DS}
  Let $[\FALSE/\D]$ be the substitution that performs $[\FALSE/\Dx{\aLoc}]$
  for every $\aLoc$.
  Let $\DS{\aLoc}{\amode}$ and $\DL{\aLoc}{\amode}$ be defined:
  \begin{align*}
    \DS{\aLoc}{\mRLX}{\bForm}&=\bForm[\TRUE/\Dx{\aLoc}] 
    &\DL{\aLoc}{\mRLX}&=\TRUE
    \\
    \DS{\aLoc}{\mRA}{\bForm}&=\bForm[\FALSE/\D]
    &\DL{\aLoc}{\mRA}&=\Dx{\aLoc}
    \\
    \DS{\aLoc}{\mSC}{\bForm}&=\bForm[\FALSE/\D]
    &\DL{\aLoc}{\mSC}&=\Dx{\aLoc}
  \end{align*}
\end{definition}

\begin{definition}[$\xCO$/$\xRASC$/$\xDGA$]
  \label{def:pomsets-down}
  Update Def.~\ref{def:pomsets-ra} to:
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qw{\aLoc}\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$.
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land((\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
  \item[\ref{L6})]
    $\aTr{\dEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land\bForm$.
  \end{enumerate}
\end{definition}

\subsection{Read-Read dependencies ($\xRRD$)}

We use the uninterpreted symbol $\RW$.

\begin{definition}[$\xRRD$]
  \label{def:pomsets-rr}
  Update Def.~\ref{def:pomsets-trans} to:
  \begin{enumerate}
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm$,
  \end{enumerate}  
\end{definition}

We repeat all of the rules for preconditions and transformers,
Only \ref{L5} and \ref{T3} are changed from Def.~\ref{def:pomsets-down}.
We repeat \ref{L4}, although it is unchanged from Def.~\ref{def:pomsets-trans}.

\begin{definition}[$\xCO$/$\xRASC$/$\xDGA$/$\xRRD$]
  \label{def:pomsets-down-rr}
  Update Def.~\ref{def:pomsets-trans} to:
  \begin{enumerate}
  \item[\ref{S3})]
    $\labelingForm(\aEv)$ implies $\QS{\aLoc}{\amode}\land\aExp{=}\aVal$,
  \item[\ref{L3})]
    $\labelingForm(\aEv)$ implies $\QL{\aLoc}{\amode}$.
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qw{\aLoc}\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$.
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies $\aVal{=}\aReg\limplies\bForm$, 
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land(\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
  \item[\ref{L6})]
    $\aTr{\dEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land\bForm$.
  \end{enumerate}  
\end{definition}

[Control dependencies into reads as in MP with release on right and control
dependency on left.]

% \begin{definition}
%   Let $[\aForm/\RW{}]$ be the substitution that replaces $\RW{}$ by $\aForm$.
% \end{definition}

\subsection{Proof Strategy}

@@

\begin{comment}
  Internal visibility:
  If $\bEv\xpoloc\aEv$ then either
  \begin{itemize}
  \item %$\bEv\xca\aEv$:
    $\bEv:\DWP{x}{1}\xco\aEv:\DWP{x}{2}$ 
  \item %$\bEv\xrfx\aEv$:
    $\bEv:\DWP{x}{1}\xrfx\aEv:\DRP{x}{1}$
  \item %$\bEv\mathrel{\xco\xrfx}\aEv$:
    $\bEv:\DWP{x}{1}\xco\DWP{x}{2}\xrfx\aEv:\DRP{x}{2}$
  \item
    $\bEv:\DRP{x}{0}\xfr\aEv:\DWP{x}{1}$\\
    $\bEv:\DRP{x}{0}\xrfxinv\DWP{x}{0}\xco\aEv:\DWP{x}{1}$
  \item %$\bEv\mathrel{\xfr\xrfx}\aEv$:
    $\bEv:\DRP{x}{0}\xfr\DWP{x}{1}\xrfx\aEv:\DRP{x}{1}$\\
    $\bEv:\DRP{x}{0}\xrfxinv\DWP{x}{0}\xco\DWP{x}{1}\xrfx\aEv:\DRP{x}{1}$
  \item %$\bEv\mathrel{\xrfxinv\xrfx}\aEv$:
    $\bEv:\DRP{x}{1}\xrfxinv\DWP{x}{1}\xrfx\aEv:\DRP{x}{1}$
    % \begin{tikzinline}
    %   \node(a){$\DWP{x}{1}$};
    %   \node(b)[right=1.5em of a]{$\DRP{x}{1}$};
    %   \node(c)[right=1.5em of b]{$\DRP{x}{1}$};
    %   \rfx{a}{b}
    %   \rfx[out=15,in=165]{a}{c}
    % \end{tikzinline}
  \end{itemize}
  If $\bEv\xpoloc\aEv$ then the following are impossible:
  \begin{itemize}
  \item $\bEv:\DWP{x}{2}\xcoinv\aEv:\DWP{x}{1}$ 
  \item $\bEv:\DWP{x}{1}\xfrinv\aEv:\DRP{x}{0}$ 
  \item $\bEv:\DRP{x}{1}\xrfxinv\aEv:\DWP{x}{1}$ 
  \item $\bEv:\DRP{x}{2}\xrfxinv\DWP{x}{2}\xcoinv\aEv:\DWP{x}{1}$
  \item $\bEv:\DRP{x}{1}\xrfxinv\DWP{x}{1}\xfrinv\aEv:\DRP{x}{0}$
  \end{itemize}
\end{comment}

\endinput

@@

An ARM execution graph $G$ defines many relations, including program order
($\rpox$), reads-from ($\rrfx$), coherence ($\rco$) and several dependency
orders.  From these are derived:
\begin{itemize}
\item ${\rpoloc}$, which is the subrelation of $\rpox$ that only relates
  actions on the same location,
\item ${\rob}$, which is required to be acyclic (\ref{external}), and
\item $\reco$, with the requirement that ${\rpoloc}\cup{\reco}$ be acyclic (\ref{sc-per-loc}).
\end{itemize}
% Let $G$ be an execution graph satisfying the ARM consistency 
% requirements.
Given an execution graph $G$, we say that $\aEv$ is an \emph{internal read} if $\aEv\in\fcodom({\rpox}\cap {\rrfx})$.

@@

% Use these to refer to the rules in text:
% \ref{rf-comp} 
% \ref{co-tot}
% \ref{sc-per-loc}
% \ref{external}

@@

From $G$ we construct a candidate pomset $\aPS$ as follows:
\begin{itemize}
\item $\Event= \textsf{E}$,
\item $\labelingAct(\aEv)=\tau \mathsf{lab}(e)$, if $\aEv$ is a relaxed
  internal read, 
\item $\labelingAct(\aEv)=\mathsf{lab}(e)$, if $\aEv$ is not a relaxed
  internal read,
\item $\labelingForm(\aEv)=\TRUE$,
  % \item ${\le} = {\rob}$, and
\item ${\gtN} = ({\rob} \cup {\reco})^*$
\end{itemize}
To reempphasize, in this candidate pomset, $\rob$ is calculated by considering the definition of $\rob$ without $\rrfi$, ie.:
\begin{align*}
  \tag{Dependency order}
  {\rdob} &\eqdef\smash{
    ({\raddr}\cup{\rdata});
    \cup ({\rctrl}\cup{\rdata}); {\IDW}; {\rcoi}^?
    \cup {\raddr}; {\rpox}; {\IDW}
  }
  \\
  \tag{Barrier order}
  {\rbob} &\eqdef\smash{
    {\IDAcq}; {\rpox}
    \cup {\rpox};{\IDRel}; {\rcoi}^?
  }
  \\
  \tag{Acyclic order}
  {\rob} &\eqdef\smash{
    ({\robs} \cup {\rdob} \cup {\rbob})^+
  }
\end{align*}
